---
title: "Analysis for Stream Microbial Metacommunity"
output: pdf_document
---
## Initial Setup
```{r}
rm(list=ls())
setwd("~/GitHub/HJA-streams/analysis")
opar <- par()
```

### Import Source Files and Packages
```{r}
source("MothurTools.R")

require("vegan")
require("png")
require("raster")
require("sp")
require("rgdal")
require("SoDA")
require("grid")
require("simba")
require("vegetarian")
require("geoR")


se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}

error.bar <- function(x, y, upper, lower=upper, length=0.1,...){
  if(length(x) != length(y) | length(y) !=length(lower) | length(lower) != length(upper))
    stop("vectors must be same length")
  arrows(x,y+upper, x, y-lower, angle=90, code=3, length=length, ...)
}

```


### Import Shared, Design, and Environment Files
```{r}
# Define Inputs
# Design = general design file for experiment
# shared = OTU table from mothur with sequence similarity clustering
# Taxonomy = Taxonomic information for each OTU
design <- "../data/design.txt"
shared <- "../data/hja_streams.final.shared"
taxon  <- "../data/hja_streams.final.0.03.taxonomy"
env    <- "../data/hja_env.csv"

# Import Design
design.total <- read.delim(design, header=T, row.names=1)

# Import Shared Files
OTUs <- read.otu(shared = shared, cutoff = "0.03")         # 97% Similarity

# Import Taxonomy
OTU.tax <- read.tax(taxonomy = taxon, format = "rdp")

# Import Env
env.total <- read.csv(env, header=T)

```

### Data Transformations
```{r}
# Remove OTUs with less than two occurances across all sites
OTUs <- OTUs[, which(colSums(OTUs) >= 2)]

# Sequencing an Good's Coverage
# Sequencing Coverage
coverage <- rowSums(OTUs)

# Good's Coverage
goods <- function(x = ""){
  1 - (sum(x == 1) / rowSums(x))
}
goods.c <- goods(OTUs)

# Remove Low Coverage Samples
lows <- which(coverage < 7000)
OTUs <- OTUs[-which(coverage < 7000), ]
design <- design.total[-which(coverage < 7000), ]
env <- env.total[-which(coverage < 7000), ]

# Remove orthogonal vectors
env <- env[c(1:11, 13, 16, 18, 19)]
env.mat <- as.matrix(env[10:15])
env.mat[51,5] <- 150
env.mat[52,5] <- 150
for(i in 1:nrow(env.mat)){
  if(env.mat[i, 5] < 0){
    env.mat[i, 5] <- 0.0001
  }
  if(env.mat[i, 6] < 0){
    env.mat[i, 6] <- 0.0001
  }
}
env.mat <- scale(env.mat)

# Distance Matrix
xy <- cbind(env$longitude, env$latitude)
geo.dists <- geoXY(env$latitude, env$longitude)
xy <- project(xy, "+proj=utm +zone=10 +ellps=WGS84")
dist.mat <- as.matrix(dist(xy, method = "euclidean"))
# 
# # Connectivity Matrix 
# adj.mat <- as.matrix(read.csv("../data/undirected-matrix.csv", header=T))
# row.names(adj.mat) <- adj.mat[,1]
# adj.mat <- adj.mat[,-1]
# adj.mat <- (adj.mat == 1) * 1
# adj.mat <- adj.mat[-which(coverage < 7000), -which(coverage < 7000)]

# Make Relative Abundence Matrices
OTUsREL <- decostand(OTUs, method = "total")

# Transform Relative Abundances
OTUsREL.log <- decostand(OTUs, method = "log")
```

## Calculate Alpha Diversity

### Richness and Rarefaction
```{r}
# Observed Richness
S.obs <- rowSums((OTUs > 0) * 1)

# Simpson's Evenness
SimpE <- function(x = ""){
  x <- as.data.frame(x)
  D <- vegan::diversity(x, "inv")
  S <- sum((x > 0) * 1) 
  E <- (D)/S 
  return(E)
}
simpsE <- round(apply(OTUsREL, 1, SimpE), 3)
shan <- vegan::diversity(OTUsREL, index = "shannon")
simpsD <- vegan::diversity(OTUsREL, index = "invsimpson")

# Rarefaction
min.N <- min(rowSums(OTUs))
S.rarefy <- rarefy(x = OTUs, sample = (min.N/2), se = TRUE)
#rarecurve(x = OTUs, step = 20, col = "blue", cex = 0.6, las = 1)
rare.d <- t(S.rarefy)
colnames(rare.d) <- c("S.rare", "se.rare")
alpha.div <- cbind(design, S.obs, simpsE, shan, simpsD, rare.d)

# Seperate data based on water and sediment samples
alpha.water <- alpha.div[alpha.div$habitat == "water",]
alpha.sed <- alpha.div[alpha.div$habitat == "sediment", ]

# Differences in diversity (richness and evenness) between habitats
t.test(alpha.water$S.rare, alpha.sed$S.rare)
mean(c(358.0118, 120.1305)) / mean(alpha.div$S.rare[alpha.div$habitat == "sediment"])

t.test(a.1$S.rare, a.2$S.rare)


```

### Numbers Equivalents
```{r}
# Numbers Equivalents
s.size <- as.vector(c(rowSums(OTUsREL)))
hja.alpha <- d(OTUsREL, lev = "alpha", q = 1, boot = T, 
               boot.arg = list(s.sizes = s.size))
hja.beta <- d(OTUsREL, lev = "beta", q = 1, boot = T, 
              boot.arg = list(s.sizes = s.size))
hja.gamma <- d(OTUsREL, lev = "gamma", q = 1, boot = T, 
               boot.arg = list(s.sizes = s.size))

water.alpha <- d(OTUsREL[which(design$habitat == "water"),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water")]))
sediment.alpha <- d(OTUsREL[which(design$habitat == "sediment"),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment")]))

water.beta <- d(OTUsREL[which(design$habitat == "water"),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water")]))
sediment.beta <- d(OTUsREL[which(design$habitat == "sediment"),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment")]))

water.alpha$D.Value / sediment.alpha$D.Value
water.beta$D.Value / sediment.beta$D.Value

# Calculate Water Alpha Diversity
water.alpha.1 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 1),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 1)]))
water.alpha.2 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 2),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 2)]))
water.alpha.3 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 3),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 3)]))
water.alpha.4 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 4),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 4)]))
water.alpha.5 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 5),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 5)]))
# Calculate Water Beta Diversity
water.beta.1 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 1),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 1)]))
water.beta.2 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 2),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 2)]))
water.beta.3 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 3),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 3)]))
water.beta.4 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 4),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 4)]))
water.beta.5 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 5),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 5)]))

# Calculate Water Gamma Diversity
water.gamma.1 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 1),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 1)]))
water.gamma.2 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 2),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 2)]))
water.gamma.3 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 3),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 3)]))
water.gamma.4 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 4),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 4)]))
water.gamma.5 <- d(OTUsREL.log[which(design$habitat == "water" & design$order == 5),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "water" & 
                                                          design$order == 5)]))

# Calculate sediment alpha diversity
sed.alpha.1 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 1),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 1)]))
sed.alpha.2 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 2),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 2)]))
sed.alpha.3 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 3),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 3)]))
sed.alpha.4 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 4),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 4)]))
sed.alpha.5 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 5),], 
                 lev = "alpha", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 5)]))

# Calculate sediment beta diversity
sed.beta.1 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 1),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 1)]))
sed.beta.2 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 2),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 2)]))
sed.beta.3 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 3),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 3)]))
sed.beta.4 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 4),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 4)]))
sed.beta.5 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 5),], 
                 lev = "beta", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 5)]))

# Calculate Sediment Gamma Diversity
sed.gamma.1 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 1),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 1)]))
sed.gamma.2 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 2),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 2)]))
sed.gamma.3 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 3),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 3)]))
sed.gamma.4 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 4),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 4)]))
sed.gamma.5 <- d(OTUsREL.log[which(design$habitat == "sediment" & design$order == 5),], 
                 lev = "gamma", q = 1, boot = T, 
                 boot.arg = list(s.sizes = s.size[which(design$habitat == "sediment" & 
                                                          design$order >= 5)]))

# Create Vectors of Diversity Measures
water.a <- c(water.alpha.1$D.Value, water.alpha.2$D.Value, water.alpha.3$D.Value, 
             water.alpha.4$D.Value, water.alpha.5$D.Value)
water.a.se <- c(water.alpha.1$StdErr, water.alpha.2$StdErr, water.alpha.3$StdErr, 
             water.alpha.4$StdErr, water.alpha.5$StdErr)
water.b <- c(water.beta.1$D.Value, water.beta.2$D.Value, water.beta.3$D.Value, 
             water.beta.4$D.Value, water.beta.5$D.Value)
water.b.se <- c(water.beta.1$StdErr, water.beta.2$StdErr, water.beta.3$StdErr, 
             water.beta.4$StdErr, water.beta.5$StdErr)
water.g <- c(water.gamma.1$D.Value, water.gamma.2$D.Value, water.gamma.3$D.Value, 
             water.gamma.4$D.Value, water.gamma.5$D.Value)
water.g.se <- c(water.gamma.1$StdErr, water.gamma.2$StdErr, water.gamma.3$StdErr, 
             water.gamma.4$StdErr, water.gamma.5$StdErr)

sed.a <- c(sed.alpha.1$D.Value, sed.alpha.2$D.Value, sed.alpha.3$D.Value, 
             sed.alpha.4$D.Value, sed.alpha.5$D.Value)
sed.a.se <- c(sed.alpha.1$StdErr, sed.alpha.2$StdErr, sed.alpha.3$StdErr, 
             sed.alpha.4$StdErr, sed.alpha.5$StdErr)
sed.b <- c(sed.beta.1$D.Value, sed.beta.2$D.Value, sed.beta.3$D.Value, 
             sed.beta.4$D.Value, sed.beta.5$D.Value)
sed.b.se <- c(sed.beta.1$StdErr, sed.beta.2$StdErr, sed.beta.3$StdErr, 
             sed.beta.4$StdErr, sed.beta.5$StdErr)
sed.g <- c(sed.gamma.1$D.Value, sed.gamma.2$D.Value, sed.gamma.3$D.Value, 
             sed.gamma.4$D.Value, sed.gamma.5$D.Value)
sed.g.se <- c(sed.gamma.1$StdErr, sed.gamma.2$StdErr, sed.gamma.3$StdErr, 
             sed.gamma.4$StdErr, sed.gamma.5$StdErr)
orders <- c(1,2,3,4,5)
water.hill <- as.data.frame(cbind(orders, water.a, water.a.se, water.b, water.b.se, water.g, water.g.se))
sed.hill <- as.data.frame(cbind(orders, sed.a, sed.a.se, sed.b, sed.b.se, sed.g, sed.g.se))

# Plot by stream order to visualize
plot(water.hill$orders, water.hill$water.a, type = "l")
plot(water.hill$orders, water.hill$water.b, type = "l")
plot(water.hill$orders, water.hill$water.g, type = "l")
plot(sed.hill$orders, sed.hill$sed.a, type = "l")
plot(sed.hill$orders, sed.hill$sed.b, type = "l")
plot(sed.hill$orders, sed.hill$sed.g, type = "l")

model.water.b <- lm(water.hill$water.b ~ water.hill$orders)
summary(model.water.b)
conf95.water.b <- predict.lm(object = model.water.b, interval="confidence")

model.sed.b <- lm(sed.hill$sed.b ~ sed.hill$orders)
summary(model.sed.b)
conf95.sed.b <- predict.lm(object = model.sed.b, interval="confidence")
```


### Figure 1: Rarefied diversity in sediment and surface-water communities
```{r}
png(filename = "../figures/Figure1.png",
    width = 1200, height = 1200, res = 96*2)
par(mar = c(5, 5, 3, 2) + 0.3)
boxplot(S.rare ~ habitat, data = alpha.div, at = c(3,1),
     names = c("Sediment", "Water"), col = c("grey", "white"),
     xlim = c(0,4),
     yaxt = "n", ylab = "", xlab = "", cex.axis = 1.2)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 1, labels = F, at = c(1,3), lwd.ticks = 2, cex.axis = 1.2, las = 1)
box(lwd = 2)
text(3, 1700, "*", cex = 2)
mtext("Habitat Type", side = 1, line = 3, cex = 1.5)
mtext("Species Richness", side = 2, line = 4, cex = 1.5)
dev.off()
graphics.off()

```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Higher richness in surface water than in sediments."}
img <- readPNG("../figures/Figure1.png")
grid.raster(img)
```



### Calculate Bray-Curtis across stream orders
```{r}
order.1.bray.w <- vegdist(OTUsREL.log[which(design$order==1 & design$habitat == "water"),], method = "bray")
order.2.bray.w <- vegdist(OTUsREL.log[which(design$order>=2 & design$habitat == "water"),], method = "bray")

order.1.bray.s <- vegdist(OTUsREL.log[which(design$order==1 & design$habitat == "sediment"),], method = "bray")
order.2.bray.s <- vegdist(OTUsREL.log[which(design$order>=2 & design$habitat == "sediment"),], method = "bray")

order.1.bray.w.list <- liste(order.1.bray.w)[,3]
order.2.bray.w.list <- liste(order.2.bray.w)[,3]

order.1.bray.s.list <- liste(order.1.bray.s)[,3]
order.2.bray.s.list <- liste(order.2.bray.s)[,3]

orders.bray.w <- as.data.frame(cbind(c(
        rep(1, length(order.1.bray.w.list)),
        rep(2, length(order.2.bray.w.list))),
      c(order.1.bray.w.list, order.2.bray.w.list)))
colnames(orders.bray.w) <- c("order", "dist")
model.orders.w <- lm(orders.bray.w$dist ~ orders.bray.w$order)
(modsum.w <- summary(model.orders.w))

orders.bray.s <- as.data.frame(cbind(c(
        rep(1, length(order.1.bray.s.list)),
        rep(2, length(order.2.bray.s.list))),
      c(order.1.bray.s.list, order.2.bray.s.list)))
colnames(orders.bray.s) <- c("order", "dist")
model.orders.s <- lm(orders.bray.s$dist ~ orders.bray.s$order)
(modsum.s <- summary(model.orders.s))

# model.orders.w <- aov(orders.bray.w$dist ~ as.factor(orders.bray.w$order))
# (modsum.w <- summary(model.orders.w))
# TukeyHSD(model.orders.w)
# 
# model.orders.s <- aov(orders.bray.s$dist ~ as.factor(orders.bray.s$order))
# (modsum.s <- summary(model.orders.s))
# TukeyHSD(model.orders.s)
```

# Figure 3: Beta-diversity is higher among headwaters than among higher order streams.
```{r}
png(filename = "../figures/Figure3.png",
    width = 1200, height = 1200, res = 96*2)
par(mar = c(5, 5, 2, 2) + 0.1)
boxplot(dist ~ order, data = orders.bray.w, at = c(1,4), xlim = c(0,6), ylim = c(0.4, 0.9), 
        col = "white", xlab = "", xaxt = "n", yaxt = "n")
boxplot(dist ~ order, data = orders.bray.s, at = c(2,5), add = T, col = "grey",
        xlab = "", xaxt = "n", yaxt = "n")
axis(side = 1, labels = F, at = c(1.5, 4.5), lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 3, labels = F, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 4, labels = F, lwd.ticks = 2, cex.axis = 1.2, las = 1)
box(lwd=2)
mtext(side = 2, cex = 1.5, line = 3.5, expression(paste(beta,"-diversity")))
mtext(side = 1, cex = 1.5, line = 3, "Stream Order")
mtext(side = 1, at = c(1.5,4.5), line = 1, c("Headwaters","Higher-order"), cex = 1.2)

legend("bottomleft", c("Water", "Sediment"),
         pt.bg = c("white", "grey"), pch = c(22,22), cex = 1.5, bty = "n")

dev.off()
graphics.off()
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Microbial Commuinity Shifts"}
img <- readPNG("../figures/Figure3.png")
grid.raster(img)
```
### Ordination

#### PCoA 
```{r}
# All HJA Catchment
hja.db <- vegdist(OTUsREL, method = "bray", upper = TRUE, diag = TRUE)
hja.pcoa <- cmdscale(hja.db, eig=TRUE, k=3)
var1 <- round(hja.pcoa$eig[1] / sum(hja.pcoa$eig),3) * 100
var2 <- round(hja.pcoa$eig[2] / sum(hja.pcoa$eig),3) * 100
var3 <- round(hja.pcoa$eig[3] / sum(hja.pcoa$eig),3) * 100

# Sediments Only
sediment <- OTUsREL[which(design$habitat == "sediment"),]
sediment.db <- vegdist(sediment, method = "bray", diag = T)
sediment.pcoa <- cmdscale(sediment.db, eig=TRUE, k=3)
sed.design <- design[which(design$habitat == "sediment"),]
s.var1 <- round(sediment.pcoa$eig[1] / sum(sediment.pcoa$eig),3) * 100
s.var2 <- round(sediment.pcoa$eig[2] / sum(sediment.pcoa$eig),3) * 100
s.var3 <- round(sediment.pcoa$eig[3] / sum(sediment.pcoa$eig),3) * 100

# Water Only
water <- OTUsREL[which(design$habitat == "water"),]
water.db <- vegdist(water, method = "bray", diag = T)
water.pcoa <- cmdscale(water.db, eig=TRUE, k=3)
water.design <- design[which(design$habitat == "water"),]
w.var1 <- round(water.pcoa$eig[1] / sum(water.pcoa$eig),3) * 100
w.var2 <- round(water.pcoa$eig[2] / sum(water.pcoa$eig),3) * 100
w.var3 <- round(water.pcoa$eig[3] / sum(water.pcoa$eig),3) * 100

# Lookout Creek Watershed Only
lc <- OTUsREL[which(design$watershed == "LC"),]
lc.db <- vegdist(lc, method = "bray", diag = T)
lc.pcoa <- cmdscale(lc.db, eig=TRUE, k=3)
lc.design <- design[which(design$watershed == "LC"),]
lc.var1 <- round(lc.pcoa$eig[1] / sum(lc.pcoa$eig),3) * 100
lc.var2 <- round(lc.pcoa$eig[2] / sum(lc.pcoa$eig),3) * 100
lc.var3 <- round(lc.pcoa$eig[3] / sum(lc.pcoa$eig),3) * 100

# Watershed 01 Only
w1 <- OTUsREL[which(design$watershed == "WS01"),]
w1.db <- vegdist(w1, method = "bray", diag = T)
w1.pcoa <- cmdscale(w1.db, eig=TRUE, k=3)
w1.design <- design[which(design$watershed == "WS01"),]
w1.var1 <- round(w1.pcoa$eig[1] / sum(w1.pcoa$eig),3) * 100
w1.var2 <- round(w1.pcoa$eig[2] / sum(w1.pcoa$eig),3) * 100
w1.var3 <- round(w1.pcoa$eig[3] / sum(w1.pcoa$eig),3) * 100
```

#### Figure 4: Sediments & Water Communities PCoA
```{r}
png(filename = "../figures/Figure4.png",
    width = 1200, height = 1200, res = 96*2)

par(mar = c(5, 5, 3, 2) + 0.1)
plot(hja.pcoa$points[ ,1], hja.pcoa$points[ ,2], ylim = c(-0.5, 0.4), xlim = c(-.35, .35),
     xlab = paste("PCoA 1 (", var1, "%)", sep = ""),
     ylab = paste("PCoA 2 (", var2, "%)", sep = ""), 
     pch = 19, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = F)
axis(side = 1, labels = T, at = c(-0.3,-.15,0,.15,.3), lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 3, labels = F, at = c(-0.3,-.15,0,.15,.3), lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 4, labels = F, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)
points(hja.pcoa$points[which(design$habitat == "sediment"),1], 
       hja.pcoa$points[which(design$habitat == "sediment"),2],
       pch=21, cex=2, bg="grey")
points(hja.pcoa$points[which(design$habitat == "water"),1], 
       hja.pcoa$points[which(design$habitat == "water"),2],
       pch=24, cex=2, bg="white")
legend("topright", c("Water", "Sediment"),
         pt.bg = c("white", "grey"), pch = c(24,21), cex = 1.5, bty = "n")

dev.off()
graphics.off()


```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Ordination across the HJ Andrews Catchment"}
img <- readPNG("../figures/Figure4.png")
grid.raster(img)
```

# Important drivers of community structure
```{r}
# hja.ano <- anosim(hja.db, grouping = design$habitat)
# summary(hja.ano)
# plot(hja.ano)

adonis(hja.db ~ design$order, permutations = 999)
adonis(hja.db ~ design$habitat * design$watershed + design$order, permutations = 999)

```

### Identify Unique OTUs across habitats
```{r}
sediment.only <- OTUs[, which(colSums(OTUs[which(design$habitat == "water"),]) == 0)]
water.only <- OTUs[, which(colSums(OTUs[which(design$habitat == "sediment"),]) == 0)]
water.sed <- OTUs[, which(colSums(OTUs[which(design$habitat == "sediment"),]) > 1 & 
                            colSums(OTUs[which(design$habitat == "water"),]) > 1)]
in.water <- OTUs[, which(colSums(OTUs[which(design$habitat == "water"),]) > 1)]
in.water.pa <- (in.water > 0) * 1
in.sediment <- OTUs[, which(colSums(OTUs[which(design$habitat == "sediment"),]) > 1)]
in.sediment.pa <- (in.sediment > 0) * 1

water.frac <- as.vector(rowSums(in.water.pa[which(design$habitat == "water"), which(colSums(in.water.pa[which(design$habitat == "sediment"),]) > 0)]) / rowSums(in.water.pa[which(design$habitat == "water"),]))
se(rowSums(in.water[which(design$habitat == "water"), which(colSums(in.water[which(design$habitat == "sediment"),]) > 1)]) / rowSums(in.water[which(design$habitat == "water"),]))

sed.frac <- as.vector(rowSums(in.sediment.pa[which(design$habitat == "sediment"), which(colSums(in.sediment.pa[which(design$habitat == "water"),]) > 1)]) / rowSums(in.sediment.pa[which(design$habitat == "sediment"),]))
se(rowSums(in.sediment[which(design$habitat == "sediment"), which(colSums(in.sediment[which(design$habitat == "water"),]) > 1)]) / rowSums(in.sediment[which(design$habitat == "sediment"),]))
```

### Figure 5: Proportion unique taxa
```{r}
png(filename = "../figures/Figure5.png",
    width = 1200, height = 1200, res = 96*2)
par(mar = c(5, 5, 2, 2) + 0.1)
boxplot((1-water.frac), (1-sed.frac), names = c("Water", "Sediments"), col = c("white", "grey"),
        at = c(1,3), xlim = c(0,4), yaxt = "n", cex.axis = 1.2)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 1, labels = F, at = c(1,3), lwd.ticks = 2, cex.axis = 1.2, las = 1)
box(lwd=2)
mtext("Proportion Unique Taxa", side = 2, line = 3.5, cex = 1.5)
mtext("Habitat Type", side = 1, line = 3, cex = 1.5)
text(x = 3, y = 0.25, "*", cex = 2)
dev.off()
graphics.off()

t.test(1-water.frac, 1-sed.frac)
mean(1 - water.frac) / mean(1 - sed.frac)
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Surface waters house more unique taxa than sediments do."}
img <- readPNG("../figures/Figure5.png")
grid.raster(img)
```

# Distance Decay Relationships

### Calculate Dendritic Distances
```{r}
require('igraph')
adj.mat <- as.matrix(read.csv("../data/undirected-matrix.csv", header=T))
row.names(adj.mat) <- adj.mat[,1]
adj.mat <- adj.mat[,-1]
adj.mat <- (adj.mat == 1) * 1

stream.network <- graph_from_adjacency_matrix(adjmatrix = adj.mat)

png(filename = "../figures/stream-network.png", 
    height = 4800, width = 4800, res = 2*96)
plot.igraph(stream.network)
dev.off()

 # Create Dist Matrix
xy.total <- cbind(env.total$longitude, env.total$latitude)
xy.total <- project(xy.total, "+proj=utm +zone=10 +ellps=WGS84")
euc.dist.mat <- as.matrix(dist(xy.total, method = "euclidean"))
rownames(euc.dist.mat) <- env.total$sample
colnames(euc.dist.mat) <- env.total$sample

# Function to find paths along the network.
dend.dist <- function(graph = "", dist.mat = ""){

  dend.dist.mat <- matrix(data = NA, nrow = length(env.total$sample), 
                          ncol = length(env.total$sample))
  rownames(dend.dist.mat) <- env.total$sample
  colnames(dend.dist.mat) <- env.total$sample
  
  for(row in rownames(dend.dist.mat)){
    for(col in colnames(dend.dist.mat)){
      path.dist <- 0
      path <- shortest_paths(graph = graph, from = row, to = col)$vpath[[1]]
      
      if(length(path) > 1){
        for(i in 1:(length(path) - 1)){
          path.dist <- path.dist + dist.mat[path[[i]]$name, path[[i+1]]$name]
        }
      }
      
      dend.dist.mat[row,col] <- path.dist
    }
  }
  
  return(dend.dist.mat)
}

dend.dist.mat <- dend.dist(graph = stream.network, dist.mat = euc.dist.mat)

#net = graph.adjacency((dend.dist.mat * adj.mat), mode = "directed",
#                      weighted = TRUE)
#plot.igraph(net)
```

#### Make distance lists
```{r}
xy <- cbind(env$longitude, env$latitude)
xy <- project(xy, "+proj=utm +zone=10 +ellps=WGS84")
dist.mat <- as.matrix(dist(xy, method = "euclidean"))
dist.mat[!lower.tri(dists)] <- NA 
dists <- dend.dist.mat[which(rownames(dend.dist.mat) %in% rownames(design)),
                       which(rownames(dend.dist.mat) %in% rownames(design))]
dists[!lower.tri(dists)] <- NA 
env.2 <- env.mat[,2:6]+2.685105 # Make all values positive

# Water Distance Decay
water.xy <- xy[which(design$habitat == "water"),]
water.den.dist <- dists[which(design$habitat == "water"),which(design$habitat == "water")]
water.geo.dist.ls <- na.omit(liste(dist.mat[which(design$habitat == "water"),
                                            which(design$habitat == "water")], 
                                   entry = "geo.dist")[,3])
water.struc.dist <- 1 - water.db
water.env.dist <- vegdist(env.2[which(design$habitat == "water"),], method = "gower", upper = F, diag = F)
water.env.dist.ls <- liste(water.env.dist, entry = "env")[,3]
water.struc.dist.ls <- liste(water.struc.dist, entry = "struc")[,3]
water.den.dist.ls <- na.omit(liste(water.den.dist, entry = "den.dist")[,3])
water.distances <- data.frame(water.den.dist.ls, water.geo.dist.ls, water.struc.dist.ls, water.env.dist.ls)
names(water.distances) <- c("den", "geo", "struc", "env")

# Sediment Distance Decay
sed.xy <- xy[which(design$habitat == "sediment"),]
sed.den.dist <- dists[which(design$habitat == "sediment"),which(design$habitat == "sediment")]
sed.geo.dist.ls <- na.omit(liste(dist.mat[which(design$habitat == "sediment"),
                                          which(design$habitat == "sediment")], 
                                 entry = "geo.dist")[,3])
sed.struc.dist <- 1 - sediment.db
sed.env.dist <- vegdist(env.2[which(design$habitat == "sediment"),], method = "gower", upper = F, diag = F)
sed.env.dist.ls <- liste(sed.env.dist, entry = "env")[,3]
sed.struc.dist.ls <- liste(sed.struc.dist, entry = "struc")[,3]
sed.den.dist.ls <- na.omit(liste(sed.den.dist, entry = "den.dist")[,3])
sed.distances <- data.frame(sed.den.dist.ls, sed.geo.dist.ls, sed.struc.dist.ls, sed.env.dist.ls)
names(sed.distances) <- c("den", "geo", "struc", "env")
```

# Distance-Decay Relationships
```{r}
# Principal Components Analysis on HJA Environment
hja.pca <- princomp(env.mat)
summary(hja.pca)
plot(hja.pca, type = "l")
biplot(hja.pca)
pc1 <- hja.pca$scores[,1]
pc2 <- hja.pca$scores[,2]
pc3 <- hja.pca$scores[,3]

cor(pc1, env.mat)
cor(pc2, env.mat)

# Catchment-Scale
env.dists <- vegdist(env.2, method = "gower")
env.dists <- liste(env.dists, entry = "env")[,3]
geo.dists <- na.omit(liste(dist.mat, entry = "geo.dist"))
#env.dists <- liste(dist(pc1), entry = "env")[,3]
den.dists <- na.omit(liste(dists, entry = "dend.dist"))
hja.dists <- cbind(liste(hja.db, entry = "comm.struc"), env.dists)
hja.dists <- cbind(hja.dists, den.dists[,3], geo.dists[,3])
names(hja.dists)[c(5,6)] <- c("den.dists", "geo.dists")

hja.env.lm <- (lm(log(1-hja.dists$comm.struc) ~ hja.dists$env.dists))
hja.den.lm <- (lm(log(1-hja.dists$comm.struc) ~ hja.dists$den.dists))
hja.geo.lm <- (lm(log(1-hja.dists$comm.struc) ~ hja.dists$geo.dists))
hja.geo_env.lm <- lm(log(1-hja.dists$comm.struc) ~ hja.dists$geo.dists * hja.dists$env.dists)

summary(hja.geo_env.lm)
summary(hja.env.lm)
summary(hja.geo.lm)
summary(hja.den.lm)

water.pca <- princomp(env.mat[which(design$habitat == "water"),2:6])
summary(water.pca)
water.pc1 <- water.pca$scores[,1]
cor(water.pc1, env.mat[which(design$habitat == "water"),2:6])
water.env.dists <- vegdist(env.2[which(design$habitat == "water"),], method = "gower")
water.env.dists <- liste(water.env.dists, entry = "env")[,3]
water.geo.dists <- na.omit(liste(dist.mat[which(design$habitat == "water"), which(design$habitat == "water")],
                                 entry = "geo.dist"))
#water.env.dists <- liste(dist(water.pc1), entry = "env")[,3]
water.dists <- cbind(liste((1 - water.db), entry = "comm.struc"), water.env.dists, water.geo.dists)
plot(water.dists$water.env.dists, log(water.dists$comm.struc))

sed.env.dists <- vegdist(env.2[which(design$habitat == "sediment"),], method = "gower")
sed.env.dists <- liste(sed.env.dists, entry = "env")[,3]
sed.pca <- princomp(env.mat[which(design$habitat == "sediment"),2:6])
summary(sed.pca)
sed.pc1 <- sed.pca$scores[,3]
cor(sed.pc1, env.mat[which(design$habitat == "sediment"),2:6])
sed.geo.dists <- na.omit(liste(dist.mat[which(design$habitat == "sediment"), which(design$habitat == "sediment")],
                                 entry = "geo.dist"))
#sed.env.dists <- liste(dist(sed.pc1), entry = "env")[,3]
sed.dists <- cbind(liste((1-sediment.db), entry = "comm.struc"), sed.env.dists, sed.geo.dists)
plot(sed.dists$sed.env.dists, log(sed.dists$comm.struc))

sed.dd <- (lm(log(sed.dists$comm.struc) ~ sed.dists$sed.env.dists))
water.dd <- (lm(log(water.dists$comm.struc) ~ water.dists$water.env.dists ))

plot(sed.dists$geo.dist, log(sed.dists$comm.struc))
plot(water.dists$geo.dist, log(water.dists$comm.struc))

summary(sed.dd)
summary(water.dd)
```



## Figure: Sed-Water DDR
```{r}
png(filename = "../figures/Figure9.png",
    width = 1200, height = 1200, res = 96*2)
par(mfrow = c(2, 1))

par(mar = c(1, 5, 3, 3) + 0.5)
plot(water.distances$geo, log(water.distances$struc), xlab="", 
     ylab = "", xaxt="n", yaxt="n")
(water.lm <- (lm(log(water.distances$struc) ~ water.distances$geo)))
summary(water.lm)
abline(water.lm, lwd = 2)
axis(side=1, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext("log Community Similarity\nSurface Water", side = 2, line = 3, cex = 1.2)

par(mar = c(4, 5, 1, 3) + 0.5)
plot(sed.distances$geo, log(sed.distances$struc), xlab="", 
     ylab = "", xaxt = "n", yaxt = "n")
(sed.lm <- (lm(log(sed.distances$struc) ~ sed.distances$geo)))
summary(sed.lm)
abline(sed.lm, lwd = 2)
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd=2)
mtext("log Community Similarity\nSediment", side = 2, line = 3, cex = 1.2)


mtext("Dendritic distance (m)", side = 1, line = 3, cex = 1.5)

dev.off()
graphics.off()

#diffslope(sed.distances$dist, sed.distances$struc, water.distances$dist, water.distances$struc)
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Distance-decay relationship."}
img <- readPNG("../figures/Figure9.png")
grid.raster(img)
```


# Community-Environmental similarity 
```{r}
png(filename = "../figures/Figure8.png",
    width = 1200, height = 1200, res = 96*2)
par(mfrow = c(2, 1) + 0.5)

par(mar = c(1, 5, 3, 3) + 0.4)
plot(water.dists$water.env.dists, 
     log(water.dists$comm.struc), xlab="", 
     ylab = "", xaxt="n", yaxt="n")
#w.env.lm <- lm(water.dists$comm.struc ~ water.dists$water.env.dists)
abline(water.dd, lty = 1, lwd = 2)
axis(side=1, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext("log Community Similarity\nSurface Water", side = 2, line = 3, cex = 1.2)

par(mar = c(4, 5, 1, 3) + 0.4)
plot(sed.dists$sed.env.dists, 
     log(sed.dists$comm.struc), xlab="", 
     ylab = "", xaxt = "n", yaxt = "n")
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd=2)
mtext("log Community Similarity\nSediment", side = 2, line = 3, cex = 1.2)
#s.env.lm <- lm(sed.dists$comm.struc ~ sed.dists$sed.env.dists)
abline(sed.dd, lwd = 2)
#mtext("Environmental Distance (PC1)", side = 1, line = 3, cex = 1.5)
mtext("Environmental Distance", side = 1, line = 3, cex = 1.5)

#mtext("Community Similarity", side = 2, line = 3, cex = 1.5)

dev.off()
graphics.off()

#diffslope(sed.distances$dist, sed.distances$struc, water.distances$dist, water.distances$struc)


```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Beta-diversity in surface waters and sediments by elevation."}
img <- readPNG("../figures/Figure8.png")
grid.raster(img)
```

# Headwaters vs. Downstream
```{r}
s1.env.dists <- vegdist(env.2[which(design$order < 2),], method = "bray")
s1.env.dists <- liste(s1.env.dists, entry = "env")[,3]
s1.geo.dists <- na.omit(liste(dist.mat[which(design$order < 2), which(design$order < 2)],
                                 entry = "geo.dist"))
s1.db <- vegdist(OTUsREL[which(design$order < 2),])
s1.dists <- cbind(liste((1 - s1.db), entry = "comm.struc"), s1.env.dists, s1.geo.dists)


s1.lm <- (lm(log(s1.dists$comm.struc) ~ s1.dists$s1.env.dists))
abline(s1.lm)
summary(s1.lm)

shi.env.dists <- vegdist(env.2[which(design$order > 1),], method = "bray")
shi.env.dists <- liste(shi.env.dists, entry = "env")[,3]
shi.geo.dists <- na.omit(liste(dist.mat[which(design$order > 1), which(design$order > 1)],
                                 entry = "geo.dist"))
shi.db <- vegdist(OTUsREL[which(design$order > 1),])
shi.dists <- cbind(liste((1 - shi.db), entry = "comm.struc"), shi.env.dists, shi.geo.dists)
plot(shi.dists$shi.env.dists, log(shi.dists$comm.struc))
shi.lm <- (lm(log(shi.dists$comm.struc) ~ shi.dists$shi.env.dists))
abline(shi.lm)
summary(shi.lm)

# Can also add  & design$habitat == "water" to compare in seds vs water
shi.env.dists <- vegdist(env.2[which(design$order > 1 & design$habitat == "sediment"),], method = "bray")
shi.env.dists <- liste(shi.env.dists, entry = "env")[,3]
shi.geo.dists <- na.omit(liste(dist.mat[which(design$order > 1 & design$habitat == "sediment"), which(design$order > 1 & design$habitat == "sediment")],
                                 entry = "geo.dist"))
shi.db <- vegdist(OTUsREL[which(design$order > 1 & design$habitat == "sediment"),])
shi.dists <- cbind(liste((1 - shi.db), entry = "comm.struc"), shi.env.dists, shi.geo.dists)
plot(shi.dists$shi.env.dists, log(shi.dists$comm.struc))
shi.lm <- (lm(log(shi.dists$comm.struc) ~ shi.dists$shi.env.dists))
abline(shi.lm)
summary(shi.lm)

shi.dd <- (lm(log(shi.dists$comm.struc) ~ shi.dists$geo.dist))
s1.dd <- (lm(log(s1.dists$comm.struc) ~ s1.dists$geo.dist))
```

```{r}

png(filename = "../figures/Figure16.png",
    width = 1600, height = 1600, res = 96*2)
par(mfcol = c(2, 2))

par(mar = c(1, 5, 3, 0) + 0.4)
plot(s1.dists$s1.env.dists, 
     log(s1.dists$comm.struc), xlab="", 
     ylab = "", xaxt="n", yaxt="n")
abline(s1.lm, lty = 1, lwd = 2)
axis(side=1, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext("log Community Similarity", side = 2, line = 3, cex = 1.2)

par(mar = c(4, 5, 1, 0) + 0.4)
plot(shi.dists$shi.env.dists, 
     log(shi.dists$comm.struc), xlab="", 
     ylab = "", xaxt = "n", yaxt = "n")
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd=2)
mtext("log Community Similarity", side = 2, line = 3, cex = 1.2)
abline(shi.lm, lwd = 2)
mtext("Environmental Distance", side = 1, line = 3, cex = 1.5)

par(mar = c(1, 1, 3, 4) + 0.4)
plot(s1.dists$geo.dist, 
     log(s1.dists$comm.struc), xlab="", 
     ylab = "", xaxt="n", yaxt="n", xlim = c(0,12000))
abline(s1.dd, lty = 1, lwd = 2)
axis(side=1, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
mtext("Headwaters", side = 4, line = 1, cex = 1.2)
box(lwd = 2)


par(mar = c(4, 1, 1, 4) + 0.4)
plot(shi.dists$geo.dist, 
     log(shi.dists$comm.struc), xlab="", 
     ylab = "", xaxt = "n", yaxt = "n", xlim = c(0,12000))
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd=2)
abline(shi.dd, lwd = 2)
mtext("Geographic Distance", side = 1, line = 3, cex = 1.5)
mtext("Higher Orders", side = 4, line = 1, cex = 1.2)


dev.off()
graphics.off()

img <- readPNG("../figures/Figure16.png")
grid.raster(img)
```

# Catchment-scale DDRs
```{r}
png(filename = "../figures/Figure10.png",
    width = 1200, height = 1200, res = 96*2)

par(mar = c(5, 5, 3, 3) + 0.4)
plot(hja.dists$env.dists, log(1 - hja.dists$comm.struc), xlab="", 
     ylab = "", xaxt="n", yaxt="n")
abline(hja.env.lm, lwd = 2)
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext("log Community Similarity", side = 2, line = 3, cex = 1.5)
mtext("Environmental Distance (PC1)", side = 1, line = 3, cex = 1.5)
dev.off()
graphics.off()

png(filename = "../figures/Figure11.png",
    width = 1200, height = 1200, res = 96*2)

par(mar = c(5, 5, 3, 3) + 0.4)
plot(hja.dists$den.dists, log(1 - hja.dists$comm.struc), xlab="", 
     ylab = "", xaxt="n", yaxt="n")
abline(hja.den.lm, lwd = 2)
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext("log Community Similarity", side = 2, line = 3, cex = 1.5)
mtext("Dendritic distance (m)", side = 1, line = 3, cex = 1.5)
dev.off()
graphics.off()

png(filename = "../figures/Figure12.png",
    width = 1200, height = 1200, res = 96*2)

par(mar = c(5, 5, 3, 3) + 0.4)
plot(hja.dists$geo.dists, log(1 - hja.dists$comm.struc), xlab="", 
     ylab = "", xaxt="n", yaxt="n")
abline(hja.geo.lm, lwd = 2)
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext("log Community Similarity", side = 2, line = 3, cex = 1.5)
mtext("Geographic Distance (m)", side = 1, line = 3, cex = 1.5)
dev.off()
graphics.off()

png(filename = "../figures/Figure13.png",
    width = 1200, height = 1200, res = 96*2)

par(mar = c(5, 5, 3, 3) + 0.4)
plot(hja.dists$geo.dists, hja.dists$env.dists, xlab="", 
     ylab = "", xaxt="n", yaxt="n")
abline(lm(hja.dists$env.dists ~ hja.dists$geo.dists), lwd = 2)
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext("Environmental Distance (PC1)", side = 2, line = 3, cex = 1.5)
mtext("Geographic Distance (m)", side = 1, line = 3, cex = 1.5)
dev.off()
graphics.off()

png(filename = "../figures/Figure14.png",
    width = 1200, height = 1200, res = 96*2)

par(mar = c(5, 5, 3, 3) + 0.4)
plot(hja.dists$den.dists, hja.dists$env.dists, xlab="", 
     ylab = "", xaxt="n", yaxt="n")
abline(lm(hja.dists$env.dists ~ hja.dists$den.dists), lwd = 2)
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext("Environmental Distance (PC1)", side = 2, line = 3, cex = 1.5)
mtext("Dendritic Distance (m)", side = 1, line = 3, cex = 1.5)
dev.off()
graphics.off()


```

### Variation Partitioning

#### Var Part - HJA Catchment
```{r}
trunc.dist <- as.matrix(dist(geo.dists)) * adj.mat
dist.pcnm <- pcnm(trunc.dist)
ordisurf(geo.dists, scores(dist.pcnm, choi=1), bubble = 4, main = "PCNM 1")
ordisurf(geo.dists, scores(dist.pcnm, choi=2), bubble = 4, main = "PCNM 2")
ordisurf(geo.dists, scores(dist.pcnm, choi=3), bubble = 4, main = "PCNM 3")
ordisplom(dist.pcnm, choices = 1:4)
rs <- rowSums(OTUsREL.log) / sum(OTUsREL.log)
pcnmw <- pcnm(trunc.dist, w = rs)

spatial.var <- vegan::cca(OTUsREL.log ~ scores(pcnmw))
env.var <- vegan::cca(OTUsREL.log ~ env.mat[,c(1:4,6)])
env.spat.var <- vegan::cca(OTUsREL.log ~ scores(pcnmw) +
                             Condition(env.mat[,c(1:4,6)]))
spat.env.var <- vegan::cca(OTUsREL.log ~ env.mat[,c(1:4,6)] +
                             Condition(scores(pcnmw)))



sum(env.var$CCA$eig)
sum(spatial.var$CCA$eig)
sum(env.spat.var$CCA$eig)
sum(spat.env.var$CCA$eig)

hja.ca <- vegan::cca(OTUsREL.log)
sum(hja.ca$CA$eig)

var.part.1 <- sum(env.var$CCA$eig) / sum(hja.ca$CA$eig)
var.part.2 <- sum(spatial.var$CCA$eig) / sum(hja.ca$CA$eig)
var.part.3 <- sum(spat.env.var$CCA$eig) / sum(hja.ca$CA$eig)
var.part.4 <- sum(env.spat.var$CCA$eig) / sum(hja.ca$CA$eig)

print(paste("HJA Pure Env. = ", var.part.1))
print(paste("HJA Pure Space. = ", var.part.2))
print(paste("HJA Spatial + Env = ", (var.part.1 - var.part.3)))
print(paste("HJA Residual Var. = ", (1- sum(var.part.1 + var.part.2))))
```

#### Var Part - Surface Water
```{r}
env.mat <- env.mat[,c(1:4,6)]
water.dist.pcnm <- pcnm(dist(xy[which(design$habitat == "water"),]))
water.rs <- rowSums(OTUsREL.log[which(design$habitat == "water"),]) / 
  sum(OTUsREL.log[which(design$habitat == "water"),])
water.pcnmw <- pcnm(dist(xy[which(design$habitat == "water"),]), w = water.rs)

water.spatial.var <- vegan::rda(OTUsREL.log[which(design$habitat == "water"),] ~ scores(water.pcnmw))
water.env.var <- vegan::rda(OTUsREL.log[which(design$habitat == "water"),] ~ env.mat[which(design$habitat == "water"),])
water.env.spat.var <- vegan::rda(OTUsREL.log[which(design$habitat == "water"),] ~
      scores(water.pcnmw) + Condition(env.mat[which(design$habitat == "water"),]))
water.spat.env.var <- vegan::rda(OTUsREL.log[which(design$habitat == "water"),] ~ 
      env.mat[which(design$habitat == "water"),] + Condition(scores(water.pcnmw)))

sum(water.env.var$CCA$eig)
sum(water.spatial.var$CCA$eig)
sum(water.env.spat.var$CCA$eig)
sum(water.spat.env.var$CCA$eig)

water.ca <- vegan::rda(OTUsREL.log[which(design$habitat == "water"),])
sum(water.ca$CA$eig)

water.var.part.1 <- sum(water.env.var$CCA$eig) / sum(water.ca$CA$eig)
water.var.part.2 <- sum(water.spatial.var$CCA$eig) / sum(water.ca$CA$eig)
water.var.part.3 <- sum(water.spat.env.var$CCA$eig) / sum(water.ca$CA$eig)
water.var.part.4 <- sum(water.env.spat.var$CCA$eig) / sum(water.ca$CA$eig)

water.table <- matrix(c(
rbind(c("Water", "Pure Env. = ", water.var.part.1)),
rbind(c("Water", "Pure Space. = ", water.var.part.2)),
rbind(c("Water", "Spatial + Env = ", (water.var.part.1 - water.var.part.3))),
rbind(c("Water", "Residual Var. = ", (1- sum(water.var.part.1 + water.var.part.2))))
), nrow = 4, byrow = T)
colnames(water.table) <- c("Habitat", "Variation Partition", "Value")

```

#### Var Part - Sediments Only
```{r}
sediment.dist.pcnm <- pcnm(dist(xy[which(design$habitat == "sediment"),]))
sediment.rs <- rowSums(OTUsREL.log[which(design$habitat == "sediment"),]) / 
  sum(OTUsREL.log[which(design$habitat == "sediment"),])
sediment.pcnmw <- pcnm(dist(xy[which(design$habitat == "sediment"),]), w = sediment.rs)

sediment.spatial.var <- vegan::rda(OTUsREL.log[which(design$habitat == "sediment"),] ~ scores(sediment.pcnmw))
sediment.env.var <- vegan::rda(OTUsREL.log[which(design$habitat == "sediment"),] ~ env.mat[which(design$habitat == "sediment"),])
sediment.env.spat.var <- vegan::rda(OTUsREL.log[which(design$habitat == "sediment"),] ~
      scores(sediment.pcnmw) + Condition(env.mat[which(design$habitat == "sediment"),]))
sediment.spat.env.var <- vegan::rda(OTUsREL.log[which(design$habitat == "sediment"),] ~ 
      env.mat[which(design$habitat == "sediment"),] + Condition(scores(sediment.pcnmw)))

sum(sediment.env.var$CCA$eig)
sum(sediment.spatial.var$CCA$eig)
sum(sediment.env.spat.var$CCA$eig)
sum(sediment.spat.env.var$CCA$eig)

sediment.ca <- vegan::rda(OTUsREL.log[which(design$habitat == "sediment"),])
sum(sediment.ca$CA$eig)

sediment.var.part.1 <- sum(sediment.env.var$CCA$eig) / sum(sediment.ca$CA$eig)
sediment.var.part.2 <- sum(sediment.spatial.var$CCA$eig) / sum(sediment.ca$CA$eig)
sediment.var.part.3 <- sum(sediment.spat.env.var$CCA$eig) / sum(sediment.ca$CA$eig)
sediment.var.part.4 <- sum(sediment.env.spat.var$CCA$eig) / sum(sediment.ca$CA$eig)

print(paste("Sediment Pure Env. = ", sediment.var.part.1))
print(paste("Sediment Pure Space. = ", sediment.var.part.2))
print(paste("Sediment Spatial + Env = ", (sediment.var.part.1 - sediment.var.part.3)))
print(paste("Sediment Residual Var. = ", (1- sum(sediment.var.part.1 + sediment.var.part.2))))

perm.sed.spa <- permutest(sediment.spatial.var, permutations = 999)
perm.sed.env <- permutest(sediment.env.var, permutations = 999)
perm.sed.env_spa <- permutest(sediment.spatial.var, permutations = 999)
perm.sed.spa_env <- permutest(sediment.spatial.var, permutations = 999)

sediment.table <- matrix(c(
rbind(c("Sediment", "Pure Env. = ", sediment.var.part.1)),
rbind(c("Sediment", "Pure Space. = ", sediment.var.part.2)),
rbind(c("Sediment", "Spatial + Env = ", (sediment.var.part.1 - sediment.var.part.3))),
rbind(c("Sediment", "Residual Var. = ", (1- sum(sediment.var.part.1 + sediment.var.part.2))))
), nrow = 4, byrow = T)
colnames(sediment.table) <- c("Habitat", "Variation Partition", "Value")


```

```{r}
habitat <- (env[,8] == "sediment") * 1
env.dummy <- as.matrix(cbind(habitat, env[,c(10, 11, 12, 13, 15)]))

d.spatial.var <- spatial.var
d.env.var <- vegan::cca(OTUsREL.log ~ env.dummy)
d.env.spat.var <- vegan::cca(OTUsREL.log ~ scores(pcnmw) +
                             Condition(env.dummy))
d.spat.env.var <- vegan::cca(OTUsREL.log ~ env.dummy +
                             Condition(scores(pcnmw)))

sum(d.env.var$CCA$eig)
sum(d.spatial.var$CCA$eig)
sum(d.env.spat.var$CCA$eig)
sum(d.spat.env.var$CCA$eig)

d.var.part.1 <- sum(d.env.var$CCA$eig) / sum(hja.ca$CA$eig)
d.var.part.2 <- sum(d.spatial.var$CCA$eig) / sum(hja.ca$CA$eig)
d.var.part.3 <- sum(d.spat.env.var$CCA$eig) / sum(hja.ca$CA$eig)
d.var.part.4 <- sum(d.env.spat.var$CCA$eig) / sum(hja.ca$CA$eig)
```



```{r}
# sediments
hja.pa.db <- vegdist(OTUs.PA[which(design$habitat == "sediment"),], method = "bray", upper = TRUE, diag = TRUE)
hja.pa.pcoa <- cmdscale(hja.pa.db, eig=TRUE, k=3)
var1.pa <- round(hja.pa.pcoa$eig[1] / sum(hja.pa.pcoa$eig),3) * 100
var2.pa <- round(hja.pa.pcoa$eig[2] / sum(hja.pa.pcoa$eig),3) * 100
var3.pa <- round(hja.pa.pcoa$eig[3] / sum(hja.pa.pcoa$eig),3) * 100

par(mar = c(5, 5, 3, 2) + 0.1)
plot(hja.pa.pcoa$points[ ,1], hja.pa.pcoa$points[ ,2],
     xlab = paste("PCoA 1 (", var1.pa, "%)", sep = ""),
     ylab = paste("PCoA 2 (", var2.pa, "%)", sep = ""), 
     pch = 19, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = F,
     main="sediment")
axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 3, labels = F, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 4, labels = F, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)
points(hja.pa.pcoa$points[which(sed.design$watershed == "LC"),1], 
       hja.pa.pcoa$points[which(sed.design$watershed == "LC"),2],
       pch=19, cex=2, bg="wheat", col="wheat")
points(hja.pa.pcoa$points[which(sed.design$watershed == "WS01"),1], 
       hja.pa.pcoa$points[which(sed.design$watershed == "WS01"),2],
       pch=17, cex=2, bg="skyblue", col="skyblue")
text(hja.pa.pcoa$points[ ,1], hja.pa.pcoa$points[ ,2],
      cex = 0.5, labels = row.names(hja.pa.pcoa$points))
legend("topleft", c("lookout", "ws01"), col = c("wheat", "skyblue"),
         pt.bg = c("wheat", "skyblue"), pch = c(19,17), cex = 1.5, bty = "n")

# water
hja.pa.db <- vegdist(OTUs.PA[which(design$habitat == "water"),], method = "bray", upper = TRUE, diag = TRUE)
hja.pa.pcoa <- cmdscale(hja.pa.db, eig=TRUE, k=3)
var1.pa <- round(hja.pa.pcoa$eig[1] / sum(hja.pa.pcoa$eig),3) * 100
var2.pa <- round(hja.pa.pcoa$eig[2] / sum(hja.pa.pcoa$eig),3) * 100
var3.pa <- round(hja.pa.pcoa$eig[3] / sum(hja.pa.pcoa$eig),3) * 100

par(mar = c(5, 5, 3, 2) + 0.1)
plot(hja.pa.pcoa$points[ ,1], hja.pa.pcoa$points[ ,2],
     xlab = paste("PCoA 1 (", var1.pa, "%)", sep = ""),
     ylab = paste("PCoA 2 (", var2.pa, "%)", sep = ""), 
     pch = 19, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = F,
     main="water")
axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 3, labels = F, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 4, labels = F, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)
points(hja.pa.pcoa$points[which(water.design$watershed == "LC"),1], 
       hja.pa.pcoa$points[which(water.design$watershed == "LC"),2],
       pch=19, cex=2, bg="wheat", col="wheat")
points(hja.pa.pcoa$points[which(water.design$watershed == "WS01"),1], 
       hja.pa.pcoa$points[which(water.design$watershed == "WS01"),2],
       pch=17, cex=2, bg="skyblue", col="skyblue")
text(hja.pa.pcoa$points[ ,1], hja.pa.pcoa$points[ ,2],
      cex = 0.5, labels = row.names(hja.pa.pcoa$points))
legend("topleft", c("lookout", "ws01"), col = c("wheat", "skyblue"),
         pt.bg = c("wheat", "skyblue"), pch = c(19,17), cex = 1.5, bty = "n")


```

# heatmaps and clustering
```{r}
jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", 
                                 "#7FFF7F", "yellow", "#FF7F00", "red", 
                                 "#7F0000"))
order <- c(design$site[design$habitat == "water"], design$site[design$habitat == "sediment"])
levelplot(as.matrix(hja.db)[order,order], aspect = "iso", col.regions = jet.colors,
          xlab = "site", ylab = "site", scales = list(cex = 0.5),
          main = "Bray-Curtis Distance")

hja.ward <- hclust(hja.db, method = "ward.D2")
par(mar = c(1,5,2,2) + 0.1)
plot(hja.ward)

require(gplots)
heatmap.2(as.matrix(hja.db), distfun = function(x) vegdist(x, method = "bray"),
          hclustfun = function(x) hclust(x, method = "ward.D2"),
          col = jet.colors(100), trace = "none", density.info = "none")
```

# Spatial autocorrelation
```{r}
var.trend <- variog(coords = xy, data = pc1, option = "bin")

png(filename = "../figures/Figure15.png", width = 1200, height = 1200, res = 96*2)
par(c(5,5,3,1) + 0.5)
plot(var.trend, type = "l", xaxt = "n", yaxt = "n",
     ylab = "", xlab = "", xlim = c(0, 12000))
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext(side = 1, line = 3, "Geographic Distance (m)", cex = 1.5)
mtext(side = 2, line = 3, "Semivariance in PC1", cex = 1.5)
abline(v = 10000, lty = "dashed")
dev.off()
graphics.off()
img <- readPNG("../figures/Figure15.png")
grid.raster(img)

var.trend <- variog(coords = xy, data = env$conductivity, option = "bin")
par(c(5,5,3,1) + 0.5)
plot(var.trend, type = "l", xaxt = "n", yaxt = "n",
     ylab = "", xlab = "", xlim = c(0, 12000))
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd = 2)
mtext(side = 1, line = 3, "Geographic Distance (m)", cex = 1.5)
mtext(side = 2, line = 3, "Semivariance in PC1", cex = 1.5)
abline(v = 10000, lty = "dashed")
dev.off()
graphics.off()
img <- readPNG("../figures/Figure15.png")
grid.raster(img)
```



```{r}
png(filename = "../figures/Figure7.png",
    width = 1200, height = 1200, res = 96*2)
par(mar = c(5,5,2,1) + 0.1)

plot(water.hill$orders, water.hill$water.b, type = "l", lty = 1, lwd = 2,
     ylim = c(1,2.5), ylab = "", xlab = "", xaxt = "n", yaxt = "n")
#matlines(water.hill$orders, conf95.water.b, lty = c(1,0,0),
#         col = c("black", "gray50", "gray50"), lwd = c(2,1,1))
points(sed.hill$orders, sed.hill$sed.b, type = "l", lty = 2, lwd = 2)
#matlines(sed.hill$orders, conf95.sed.b, lty = c(1,0,0),
#         col = c("black", "gray50", "gray50"), lwd = c(2,1,1))
#points(water.hill$orders, (water.hill$water.b + water.hill$water.b.se), type = "l", lty = 3)
#points(water.hill$orders, (water.hill$water.b - water.hill$water.b.se), type = "l", lty = 3)
#points(sed.hill$orders, (sed.hill$sed.b + sed.hill$sed.b.se), type = "l", lty = 3)
#points(sed.hill$orders, (sed.hill$sed.b - sed.hill$sed.b.se), type = "l", lty = 3)
axis(side=1, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=2, labels=T, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=3, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
axis(side=4, labels=F, lwd.ticks=2, cex.axis=1.2, las=1)
box(lwd=2)
mtext(side = 2, cex = 1.5, line = 3, expression(paste(beta,"-diversity")))
mtext(side = 1, cex = 1.5, line = 3, "Stream Order")

legend(1, 1.35, c("Bacterioplankton", "Sediment-associated"), 
       lty = c(1,2), lwd = c(2,2), bty = "n", cex = 1.5)

dev.off()
graphics.off()

img <- readPNG("../figures/Figure7.png")
grid.raster(img)
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Beta-diversity in surface waters and sediments by stream order"}
img <- readPNG("../figures/Figure7.png")
grid.raster(img)
```

# AEM 
```{r}
site.by.edges.lc <- as.matrix(read.csv("../data/sites-by-edges_LC.csv", header=T))
row.names(site.by.edges.lc) <- site.by.edges.lc[,1]
site.by.edges.lc <- site.by.edges.lc[,-1]
site.by.edges.lc <- (site.by.edges.lc == 1) * 1

site.by.edges.w1 <- as.matrix(read.csv("../data/sites-by-edges_WS1.csv", header=T))
row.names(site.by.edges.w1) <- site.by.edges.w1[,1]
site.by.edges.w1 <- site.by.edges.w1[,-1]
site.by.edges.w1 <- (site.by.edges.w1 == 1) * 1

w1.dists <- c(1,
              1,
              255,
              10,
              4,
              6.46,
              201.47,
              196.598,
              3,
              4,
              100.3,
              116.85,
              802.3,
              1065,
              40.39,
              1)


w1.F.mat <- princomp(site.by.edges.w1)
summary(w1.F.mat)
plot(w1.F.mat)

w1.env <- princomp(env.mat[which(design$watershed == "WS01"),])
summary(w1.env)
plot(w1.env)


```

