---
title: "Analysis for Stream Microbial Metacommunity"
output: pdf_document
---

```{r}
setwd("~/GitHub/HJA-streams")
source("./analysis/InitialSetup.R", echo = TRUE)
```

## Diversity Analysis
```{r}
source("./DiversityAnalysis.R", echo = TRUE)
```

### Figure 1: Rarefied diversity in sediment and surface-water communities
```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Higher richness in surface water than in sediments."}
img <- readPNG("../figures/AlphaDiv.png")
grid.raster(img)
```


### Figure 3: Beta-diversity is higher among headwaters than among higher order streams.
```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Microbial Commuinity Shifts"}
img <- readPNG("../figures/BetaDivByOrder.png")
grid.raster(img)
```

### Ordination
```{r}
source("./Ordination.R", echo = TRUE)
```

#### Figure 4: Sediments & Water Communities PCoA
```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Ordination across the HJ Andrews Catchment"}
img <- readPNG("../figures/HJA_PCoA.png")
grid.raster(img)
```


### Identify Unique OTUs across habitats
```{r}
source("./UniqueTaxa.R", echo = TRUE)
```

### Figure 5: Proportion unique taxa
```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Surface waters house more unique taxa than sediments do."}
img <- readPNG("../figures/UniqueTaxa.png")
grid.raster(img)
```

# Distance Decay Relationships
```{r}
source("./DistanceCalcs.R", echo = T)
source("./DDRs.R", echo = T)
```

## Figure: Sed vs Water DDRs
```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Distance-decay relationship."}
img <- readPNG("../figures/DDR_WaterSed.png")
grid.raster(img)
```

# Figure: Headwaters vs Mainstem DDRs 
```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Beta-diversity in surface waters and sediments by elevation."}
img <- readPNG("../figures/DDR_HeadwaterMainstem.png")
grid.raster(img)
```








